---
title: "Essentials of Data Analytics - Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Overview
```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(caret))
suppressMessages(library(reshape2))
suppressMessages(library(broom))
suppressMessages(library(randomForest))
suppressMessages(library(performanceEstimation))
suppressMessages(library(regclass))
suppressMessages(library(GGally))
suppressMessages(library(pROC))
suppressMessages(library(plotROC))
suppressMessages(library(cowplot))
suppressMessages(library(grid))
suppressMessages(library(gridExtra))
suppressMessages(library(formattable))
suppressMessages(library(scales))
theme_set(theme_minimal())
options(warn=-1)

telco <- read_csv("Telco-Customer-Churn.csv", show_col_types = FALSE) %>%
  mutate_if(is.character, factor)

telco$SeniorCitizen <- as.factor(telco$SeniorCitizen)

telco <- telco %>% 
  select(-customerID) %>% 
  mutate_at(7, ~as.factor(case_when(. == "No phone service" ~ "No",
                                    . == "No" ~ "No", TRUE ~ "Yes"))) %>% 
  mutate_at(8, ~as.factor(case_when(. == "Fiber optic" ~ "FiberOptic",
                                    . == "DSL" ~ "DSL", TRUE ~ "No"))) %>% 
  mutate_at(c(9:14), ~as.factor(case_when(. == "No internet service" ~ "No", 
                                          . == "No" ~ "No", TRUE ~ "Yes"))) %>%
  mutate_at(17, ~as.factor(case_when(. == "Bank transfer (automatic)" ~ "BankTransferAuto", 
                                     . == "Credit card (automatic)" ~ "CreditCardAuto", 
                                     . == "Electronic check" ~ "ECheck", TRUE ~ "MailedCheck"))) %>% 
  na.omit()

telco %>% 
  group_by(gender) %>% 
  rename("Gender" = gender) %>% 
  summarise("Number of Observations" = n(),
            "Average Tenure, in months" = round(mean(tenure), 0),
            "Monthly Charges" = round(mean(MonthlyCharges), 2))
```

### Exploratory Data Analysis

What is a customer's average tenure with Telco and what are their average charges?
```{r, fig.height = 7}
t2 <- telco %>% 
  mutate(Churn2 = as.factor(ifelse(Churn == "Yes", "Former Customers", "Current Customers"))) 

g1 <- ggplot(t2, aes(x = fct_rev(Churn2), y = tenure, fill = fct_rev(Churn2))) +
  geom_bar(stat = "summary", fun = "mean", alpha = 0.6, color = "grey20", show.legend = F) +
  stat_summary(aes(label = paste(round(..y.., 0), "months")), fun = mean, geom = "text", size = 3, vjust = -0.5) +
  labs(title = "Average Customer Tenure \n", x = "", y = "Customer Tenure\n") +
  theme(plot.title = element_text(hjust = 0.5))

g2 <- ggplot(t2, aes(x = fct_rev(Churn2), y = MonthlyCharges, fill = fct_rev(Churn2))) +
  geom_bar(stat = "summary", fun = "mean", alpha = 0.6, color = "grey20", show.legend = F) +
  stat_summary(aes(label = dollar(..y..)), fun = mean, 
               geom = "text", size = 3, vjust = -0.5) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Monthly Charges \n", x = "", y = "Monthly Charges \n") +
  theme(plot.title = element_text(hjust = 0.5))

g3 <- ggplot(t2, aes(x = Contract, y = MonthlyCharges, fill = fct_rev(Churn2))) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean", alpha = 0.6, color = "grey20") +
  stat_summary(aes(label = dollar(..y..)), fun = mean, 
               geom = "text", size = 3, vjust = -0.5,
               position = position_dodge(width = 0.9)) +
  coord_cartesian(ylim = c(0, 95)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "\nAverage Monthly Charges by Contract Type", x = "\n Contract Type", 
       y = "Monthly Charges \n", fill = "") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "top", legend.justification = "left")

#options(repr.plot.width=10, repr.plot.height = 14)
grid.arrange(g1, g2, g3, ncol = 2, nrow = 2, layout_matrix = rbind(c(1,2), c(3,3)))
```

What type of account services do customers have?
```{r, fig.height=7}
g1 <- ggplot(t2, aes(x = Contract, group = fct_rev(Churn2))) + 
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat = "count",
           alpha = 0.6, color = "grey20", show.legend = F) +
  geom_text(aes(label = percent(..prop..), y = ..prop.. ), 
            size = 4, stat = "count", vjust = -0.5) +
  facet_grid(~fct_rev(Churn2)) +
  scale_y_continuous(labels = percent_format()) +
  coord_cartesian(ylim = c(0, .95)) +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Customer Churn by Contract Type\n", x = "\n Contract Type", y = "") +
  theme(plot.title = element_text(hjust = 0.5))

g2 <- ggplot(t2, aes(x = InternetService, group = fct_rev(Churn2))) + 
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat = "count",
           alpha = 0.6, color = "grey20", show.legend = F) +
  geom_text(aes(label = percent(..prop..), y = ..prop.. ), 
            size = 4, stat = "count", vjust = -0.5) +
  facet_grid(~fct_rev(Churn2)) +
  scale_y_continuous(labels = percent_format()) +
  coord_cartesian(ylim = c(0, .9)) +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "\n Customer Churn by Internet Service \n", x = "\n Internet Service", y = "") +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(g1, g2, ncol = 1)
```
<br>
Customer Attrition Demographics
```{r}
g1 <- ggplot(t2, aes(x = fct_rev(ifelse(SeniorCitizen==1, "Yes", "No")), group = Churn2)) + 
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat = "count",
           alpha = 0.6, color = "grey20", show.legend = F) +
  geom_text(aes(label = percent(..prop.., accuracy = 0.1), y = ..prop..), 
            size = 4, stat = "count", position = position_stack(vjust = 0.5)) +
  facet_grid(~fct_rev(Churn2)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  coord_cartesian(ylim = c(0, .9)) +
  labs(x = "\n Senior Citizen", y = "")

g2 <- ggplot(t2, aes(x = gender, group = Churn2)) + 
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat = "count",
           alpha = 0.6, color = "grey20", show.legend = F) +
  geom_text(aes(label = percent(..prop.., accuracy = 0.1), y = ..prop..), 
            size = 4, stat = "count", position = position_stack(vjust = 0.5)) +
  facet_grid(~fct_rev(Churn2)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  coord_cartesian(ylim = c(0, .6)) +
  labs(x = "\n Gender", y = "")

options(repr.plot.width=12, repr.plot.height=7)
grid.arrange(g1, g2, nrow = 1, top = textGrob("Customer Attrition Demographics \n",
                                              gp = gpar(fontsize = 14)))
```
 
Distributions and Correlations
```{r}
options(repr.plot.width=12, repr.plot.height=10)
telco %>% 
  select(tenure, MonthlyCharges, TotalCharges, Churn) %>%
  ggpairs(aes(color = fct_rev(Churn)), title = "Customer Account Distributions and Correlations \n",
          columnLabels = c("Tenure", "Monthly Charges", "Total Charges", "Churn"),
          upper = list(combo = wrap("box_no_facet", alpha = 0.7)),
          diag = list(continuous = wrap("densityDiag", alpha = 0.6), 
                      discrete = wrap("barDiag", alpha = 0.7, color = "grey30")),
          lower = list(combo = wrap("box_no_facet", alpha = 0.7), continuous = wrap("smooth", alpha = 0.15)))
```


### Data Preprocessing
```{r}
telco <- telco %>% 
  mutate_at(15, ~as.factor(case_when(. == "One year" ~ "OneYear", 
                                     . == "Two year" ~ "TwoYear", 
                                     TRUE ~ "Month-to-month"))) 

set.seed(1)
ind <- createDataPartition(telco$Churn, p = 0.7, list = F)
telco.train <- telco[ind,]
telco.test <- telco[-ind,]
train.resamp <- smote(Churn ~ ., data = data.frame(telco.train), perc.over = 1, perc.under = 2)

g1 <- ggplot(t2, aes(x = fct_rev(Churn2), fill = fct_rev(Churn2))) +
  geom_bar(alpha = 0.6, color = "grey30", show.legend = F) + 
  geom_text(stat = "count", size = 3.5, 
            aes(label = paste("n = ", formatC(..count.., big.mark = ","))), vjust = -0.5) +
  scale_y_continuous(labels = comma_format()) +
  labs(subtitle = "Before Resampling\n", x = "", y = "Number of Customers\n")

g2 <- ggplot(train.resamp, aes(x = fct_rev(ifelse(Churn == "Yes", "Former Customers", "Current Customers")), 
                               fill = fct_rev(Churn))) +
  geom_bar(alpha = 0.6, color = "grey30", show.legend = F) + 
  geom_text(stat = "count", size = 3.5, 
            aes(label = paste("n = ", formatC(..count.., big.mark = ","))), vjust = -0.5) +
  scale_y_continuous(labels = comma_format()) +
  labs(subtitle = "After Resampling\n", x = "", y = "")

options(repr.plot.width=9, repr.plot.height=7)
grid.arrange(g1, g2, nrow = 1, top = textGrob("Distribution of Customer Churn\n",
                                              gp = gpar(fontsize = 14)))
```

### Feature Selection
Chi-Squared Tests
```{r}
cat.var <- telco[,sapply(telco, is.factor)]
chi <- lapply(cat.var[,-17], function(x) chisq.test(cat.var[,17], x))
do.call(rbind, lapply(chi, tidy)) %>%
  arrange(p.value) %>%
  mutate_at(c(1,2), funs(round(., 3)))
```


### Predicting Customer Churn
#### 1. Logistic Regression
```{r}
set.seed(21)
ctrl <- trainControl(method = "cv", number = 10, classProbs = TRUE, 
                     summaryFunction = twoClassSummary)
glm.fit <- train(Churn ~ tenure + MonthlyCharges + InternetService + PaymentMethod + 
                 Contract + OnlineSecurity + TechSupport + PaperlessBilling, 
                 data = train.resamp, method = "glm", metric = "ROC",
                 preProcess = c("center", "scale"), trControl = ctrl)
glm.preds <- glm.fit %>% predict(telco.test)
glm.cm <- data.frame(Logistic=confusionMatrix(glm.preds, telco.test$Churn, 
                                              positive = "Yes", mode = "everything")$byClass)
confusionMatrix(glm.preds, telco.test$Churn, positive = "Yes", mode = "everything")
```

Multicollinearity
```{r}
data.frame("VIF" = VIF(glm.fit$finalModel)) %>% 
  arrange(desc(VIF))
```


### 2.Support Vector Machine
```{r}
svm.fit <- train(Churn ~ tenure + MonthlyCharges + InternetService + PaymentMethod + 
                 Contract + OnlineSecurity + TechSupport + PaperlessBilling, 
                 data = train.resamp, method = "svmLinear", metric = "ROC",
                 preProcess = c("center","scale"), trControl = ctrl)
svm.preds <- svm.fit %>% predict(telco.test)
svm.cm <- data.frame(SVM=confusionMatrix(svm.preds, telco.test$Churn, 
                                         positive = "Yes", mode = "everything")$byClass)
confusionMatrix(svm.preds, telco.test$Churn, positive = "Yes", mode = "everything")
```

### 3.Random Forest
```{r}
rf.fit <- train(Churn ~ tenure + MonthlyCharges + InternetService + PaymentMethod + 
                Contract + OnlineSecurity + TechSupport + PaperlessBilling,
                data = train.resamp, method = "rf", metric = "ROC",
                preProcess = c("center","scale"), trControl = ctrl)
rf.preds <- rf.fit %>% 
  predict(telco.test)
rf.cm <- data.frame(rf=confusionMatrix(rf.preds, telco.test$Churn, 
                                       positive = "Yes", mode = "everything")$byClass)
confusionMatrix(rf.preds, telco.test$Churn, positive = "Yes", mode = "everything")
```

### Model Evaluation and ROC Curves

1. Model Performance on the Test Set
```{r}
res.cm <- data.frame(glm.cm, svm.cm, rf.cm) %>% 
  rename("Random Forest" = rf) 
res <- data.frame(t(res.cm))
rownames(res) <- colnames(res.cm)
colnames(res) <- rownames(res.cm)
res[,c(7,6,2,11)] %>% 
  arrange(desc(F1)) %>% 
  mutate_all(percent_format(accuracy = 0.1))
```

2. ROC Curves
```{r}
Logistic <- predict(glm.fit, telco.test, type = "prob")[,2]
SVM <- predict(svm.fit, telco.test, type = "prob")[,2]
RandomForest <- predict(rf.fit, telco.test, type = "prob")[,2]

roc.data <- cbind(telco.test[,20], Logistic, SVM, RandomForest)
roc.long <- melt_roc(roc.data, d = "Churn", m = c("Logistic", "SVM", "RandomForest"))

rocplot <- ggplot(roc.long, aes(d = ifelse(D == "Yes", 1, 0), m = M, color = name)) +
  geom_roc(n.cuts = 0) + 
  style_roc(xlab = "\nFalse Positive Rate (1 - Specificity)", 
            ylab = "True Positive Rate (Sensitivity)\n") +
  labs(title = "ROC Curve Comparison on the Test Set", color = "Model") +
  theme(plot.title = element_text(hjust = 0.5))

rocplot +
  geom_abline(size = 0.5, color = "grey30") +
  annotate("text",x=.75,y=.35,label=paste("AUC of Logistic =", round(calc_auc(rocplot)$AUC[1],3))) +
  annotate("text",x=.75,y=.28,label=paste("AUC of SVM =", round(calc_auc(rocplot)$AUC[2],3))) +
  annotate("text",x=.75,y=.21,label=paste("AUC of Random Forest =", round(calc_auc(rocplot)$AUC[3],3))) +
  scale_color_discrete(breaks = c("Logistic","SVM", "RandomForest"))
```

Key Findings

```{r}
glm.fit <- train(Churn ~ tenure + MonthlyCharges + InternetService + PaymentMethod + 
                 Contract + OnlineSecurity + TechSupport + PaperlessBilling, 
                 data = telco, method = "glm", 
                 preProcess = c("center", "scale"), 
                 trControl = trainControl(method = "cv", number = 10))
summary(glm.fit$finalModel)
```


```{r}
OR <- coef(glm.fit$finalModel) %>% exp() %>% round(digits = 2) %>% as.data.frame() %>% slice(-c(1,6,8))
data.frame(Predictor = c("Tenure", "MonthlyCharges", "InternetServiceFiberOptic", 
                         "InternetServiceNo", "PaymentMethodECheck", "ContractOneYear", "ContractTwoYear",
                         "OnlineSecurityYes", "TechSupportYes", "PaperlessBillingYes"),
           OddsRatio = OR[,1],
Interpretation = c("A one month increase in tenure decreases the risk of churning by about 53%.","For every $1 increase in monthly charges, we expect to see an increase in the odds of churning by a factor of 1.39 or by 39%.","Customers with fiber optic internet are 31% more likely to churn than those with DSL.", "Those without internet are 28% less likely to churn than customers with DSL internet.", "Customers who pay with electronic checks are more likely to churn by a factor of 1.19 or by 19% compared to customers who use automatic bank transfers.", "Customers on one-year contracts are 25% less likely to churn than customers on month-to-month contracts.", "Customers on two-year contracts are 44% less likely to churn compared to those on month - to - month contracts.", "Customers with online security are 19% less likely to churn than customers without online security.", "Customers with tech support are about 17% less likely to churn than customers without tech support.", "Customers with paperless billing are 21% more likely to churn than customers without paperless billing.")) %>% 
  arrange(desc(OddsRatio))
```